---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Step 1 Result:
```{R}
library(classdata)
library(tidyverse)
ames = ames
```

```{r}
head(ames)
```

```{r}
names(ames)
```


```{r}
glimpse(ames)
```


```{r}
glimpse(ames)
```


```{r}
str(ames)
```

As a team, we inspected the ames dataset using glimpse(), names(), and str(). We found that the dataset contains 16 variables: Parcel ID, Address, Style, Occupancy, Sale Date, Sale Price, Multi Sale, YearBuilt, Acres, TotalLivingArea (sf), Bedrooms, FinishedBsmtArea (sf), LotArea(sf), AC, FirePlace, and Neighborhood. These variables include character variables such as identifiers and labels, factor variables such as housing style and neighborhood, one date variable (Sale Date), and several numeric variables such as Sale Price, YearBuilt, lot size, and living area. Based on the variable names and sample values, the dataset records property sale information, structural features, lot size, and location details for residential properties in Ames. We would expect identifiers and addresses to mostly appear as text, categorical variables to come from a limited set of repeated categories, sale dates to be valid dates, and numeric variables such as sale price, acreage, living area, and lot area to be nonnegative values. We also noticed that some variables contain missing values (NA), which may need to be considered during later analysis.





Step 2 Result:

```{r}
head(ames$`Sale Price`)
```
Sale Price is the main variable in the dataset. It's the main variable we want to predict with given unseen data. Most other variables in the dataset will be directly correlated with the sale price of the property. 








Step 3 Result:
```{R}
price <-ames$`Sale Price`
range(ames[["Sale Price"]], na.rm = FALSE)
mean(ames$`Sale Price`)
median(ames$`Sale Price`)

library(ggplot2)
#Graph with outliers
ggplot(ames, aes(x = price)) +
  geom_histogram(binwidth = 1000000, color = "blue", fill = "white") +
  labs(title = "Sales Price",
       x = "Price ($)",
       y = "Frequency")

#Graph without outliers
Q1 <- quantile(ames$`Sale Price`, 0.25, na.rm = TRUE)
Q3 <- quantile(ames$`Sale Price`, 0.75, na.rm = TRUE)
IQR_value <- IQR(ames$`Sale Price`, na.rm = TRUE)

lower_bound <- Q1 - 1.5 * IQR_value
upper_bound <- Q3 + 1.5 * IQR_value

ames_no_outliers <- ames[
  ames$`Sale Price` >= lower_bound &
  ames$`Sale Price` <= upper_bound,
]

ggplot(ames_no_outliers, aes(x = `Sale Price`)) +
  geom_histogram(binwidth = 100000, color = "blue", fill = "white") +
  labs(title = "Sales Price",
       x = "Price ($)",
       y = "Frequency")
```
The general pattern of the sale price graph is it is skewed to the right. That means majority of the houses sold are going to be in the cheaper range. There are also a few severe outliers. The range of this dataset consists of the most expensive sale being 20.5 million and the lowest sale being 0. The average is is 1,017,479. The median being 170,900.









Step 4 Result:
Nicolas Result: acres







Nicole Result:

```{r}
unique(ames$Bedrooms)
```


```{r}
nicolePlot1 <- ggplot(aes(x= Bedrooms), data = ames) +
  geom_bar(na.rm = T) +
  geom_text(stat = 'count', aes(label = after_stat(count), vjust = -.5)) +
  labs(
    title = 'Number of Bedrooms',
    x = "Number of Bedrooms")
nicolePlot1
```

The distribution of the number of bedrooms is approximately bell shaped with the mode being 3 bedrooms. There are 39 houses with 0 bedrooms. The maximum number of bedrooms is 10. 

```{r fig.height = 7}
filteredAmes <- ames[ames$`Sale Price`<= 1000000 & !is.na(ames$Bedrooms) & ames$`Sale Price`!= 0, ]

nicolePlot2 <- ggplot(filteredAmes, aes(x = factor(Bedrooms), y = `Sale Price`)) +
  geom_boxplot(na.rm = TRUE) +
  coord_cartesian(ylim = c(0, 600000)) +
  labs(
    x = 'Number of Bedrooms',
    y = 'Sale Price (USD)',
    title = 'Sale Price Boxplots: Number of Bedrooms'
  )
nicolePlot2
```
This diagram shows that the median sale prices of Ames houses generally increase with the number of bedrooms. However, houses with 6 or 7 bedrooms had a lower median sale price than houses with 5 or 8 bedrooms. 8 bedroom houses had the highest median sale price and the largest IQR of sale prices. Houses with 1, 2, 3, or 7 bedrooms all showed fairly small IQRs. There are several outliers on the high end (sale price >= $1,000,000) that are not displayed in this chart for visualization purposes.

```{r}
amesOutliers <- ames[ames$`Sale Price` >= 1000000, ]
nicolePlot3 <- ggplot(amesOutliers, aes(x = Bedrooms, y = `Sale Price`)) +
  geom_point() +
  labs(
    title = 'Ames Housing Sale Prices (>= $1,000,000) vs Number of Bedrooms',
    xlab = 'Number of Bedrooms',
    ylab = 'Sale Price (USD)',
  )
nicolePlot3
```

There does not appear to be a relationship between the number of bedrooms and the sale price of houses that sold for over $1,000,000.








Scott Result: Finished Basement Area

```{r}
summary(ames$`FinishedBsmtArea (sf)`)
```

```{r}
hist(ames$`FinishedBsmtArea (sf)`,
     main = "Histogram of Finished Basement Area",
     xlab = "Finished Basement Area (sq ft)",
     col = "lightblue",
     border = "black")
```

```{r}
boxplot(ames$`FinishedBsmtArea (sf)`,
        main = "Boxplot of Finished Basement Area",
        horizontal = TRUE,
        col = "lightgreen")
```

```{r}
plot(ames$`FinishedBsmtArea (sf)`,
     ames$`Sale Price`,
     main = "Finished Basement Area vs Sale Price",
     xlab = "Finished Basement Area (sq ft)",
     ylab = "Sale Price",
     pch = 19,
     col = "steelblue")
```
The variable Finished Basement Area (sf) measures the square footage of finished basement space in each home. The histogram shows that the distribution is right-skewed, with many homes having smaller finished basement areas and fewer homes having very large finished basements. The boxplot indicates the presence of some high-value outliers. The scatterplot comparing finished basement area and sale price suggests a positive relationship: homes with larger finished basement areas tend to have higher sale prices. However, the relationship does not appear perfectly linear, and other factors likely also influence sale price.
```{r}
cor(ames$`FinishedBsmtArea (sf)`,
    ames$`Sale Price`,
    use = "complete.obs")
```
This value shows that there is not a strong correlation between having a finished basement and the sale price increasing.
```{r}
ames$BasementFinished <- ifelse(
  ames$`FinishedBsmtArea (sf)` > 0,
  "Finished",
  "Not Finished"
)
```

```{r}
library(ggplot2)

ggplot(ames, aes(x = BasementFinished)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Finished vs Not Finished Basements",
       x = "Basement Status",
       y = "Number of Homes") +
  theme_minimal()
```
The bar chart shows the number of homes with finished versus unfinished basements. Most homes in the dataset have a finished basement, while a smaller portion have no finished basement area. This suggests that finished basements are common among properties in Ames and may contribute to differences in sale price.







Shashank Result: YearBuilt


Range of the Variable / Description of the Pattern
```{r}
summary(ames$YearBuilt)
```


```{r}
range(ames$YearBuilt, na.rm = TRUE)
```




YearBuilt (solo) Plot
```{r}
ggplot(ames, aes(x = YearBuilt)) +
  geom_histogram(bins = 30) +
  labs(
    title = "Distribution of Year Built",
    x = "Year Built",
    y = "Count"
  )
```





YearBuilt vs. Sale Price
```{r}
ggplot(ames, aes(x = YearBuilt, y = `Sale Price`)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Sale Price vs Year Built",
    x = "Year Built",
    y = "Sale Price"
  )
```






Does YearBuilt Describe any Oddities Discovered in 3?
```{r}
ames %>%
  filter(`Sale Price` == 0) %>%
  select(`Sale Price`, YearBuilt, Address, `Multi Sale`, Occupancy, Neighborhood) %>%
  head(15)
```


For my individual investigation, I chose YearBuilt as a variable that might be related to Sale Price. Based on the summary statistics, YearBuilt ranges from 0 to 2022, with a median of 1978 and a mean of 1976. Most homes in the dataset seem to have been built between the mid-1900s and early 2000s, since the first quartile is 1956 and the third quartile is 2002. The histogram shows that a large portion of the homes were built in more recent decades, especially from around the 1950s onward, while there are fewer very old homes. One thing that stands out, though, is that the minimum value is 0, which is clearly not a real construction year. That likely means some records are missing or entered incorrectly, so YearBuilt = 0 should be treated as an odd value rather than a real year.

To look at the relationship with the main variable, I made a scatterplot of Sale Price versus YearBuilt. In general, there seems to be a slight positive relationship, where newer homes tend to sell for more than older homes. At the same time, the relationship is not super strong, since there is still a lot of spread in sale prices across many different build years. The plot also has a few extremely large sale price outliers, which make most of the other points bunch together near the bottom of the graph. Because of that, the overall trend is harder to see clearly, but it still looks like newer homes are often associated with higher prices.

I also checked whether YearBuilt helps explain the oddity from Step 3, which was that some homes have Sale Price = 0. Looking at those rows, the zero-price homes were built across a wide range of years, including 1950, 1951, 1955, 1960, 1966, 1972, 1975, 1992, 1995, and 1997. Since those zero-price sales are not all concentrated in one time period, YearBuilt does not seem to explain that oddity very well. In other words, the homes with zero sale price are not all especially old or especially new. That makes it more likely that the zero values are due to something else, such as unusual transaction records or missing/placeholder sale data.



Pablo Result:

```{r}
ames_clean <- ames |>
  group_by(`Multi Sale`, `Sale Price`) |>
  mutate(
    units_in_sale = n(),
    adjusted_price = ifelse(
      !is.na(`Multi Sale`),
      `Sale Price` / units_in_sale,
      `Sale Price`
    )
  ) |>
  ungroup()


ames_clean <- ames_clean |>
  filter(
    !is.na(`TotalLivingArea (sf)`),
    !is.na(`Sale Price`),
  )

ames_clean <- ames_clean |>
  filter(
    adjusted_price > 50,
    `TotalLivingArea (sf)` > 0,  
  )

view(ames_clean)
```

On first glance, the dataset seemed to have many outliers with extraordinarily high sale prices for the given property. On further inspection, however, it was determined that entries in this dataset that were bought as a package are still listed individually, only with the Multi Sale flag set to 'Y'. As a result, the package sales with many properties being bought at once listed each property as having the sale price of the entire package. This skews the median sale price higher and makes data analysis and visualization difficult. To remedy this, I mutated the dataset to append an adjusted_price variable that divides the total sale price by the number of units sold to get an average property sale price for each unit. This only affects entries with a non null Multi Sale feature. Additionally, as part of the cleaning process, I filtered out entries with null / nonzero TotalLivingArea or null Sale Price and only saved entries with an adjusted_price of more than 50, to avoid questionable outliers. 

```{r}
ggplot(data = ames_clean, aes(x = `TotalLivingArea (sf)`)) +
  geom_histogram(binwidth = 100)
```
The result of the data cleaning is a satisfactory right skewed histogram that shows most properties as having between 1000 and 2000 square feet of total living area. 
```{r}
# priceHist <- filter(ames_clean, `adjusted_price` > 900)
ggplot(data = ames, aes(x = `Sale Price`)) +
  geom_histogram(binwidth = 10000)

ggplot(data = ames_clean, aes(x = `adjusted_price`)) +
  geom_histogram(binwidth = 10000)

outlier <- ames_clean
  # filter(`adjusted_price` < 50000)
outlier |>
  select(Address, Style, `Sale Date`, `Sale Price`, `Multi Sale`, `adjusted_price`) |>
  arrange(`adjusted_price`) |>
  view()

multi_sales_collapsed <- ames_clean |>
  filter(`Multi Sale` == "Y") |>
  group_by(`Sale Date`, `Sale Price`) |>
  summarise(
    address_example = first(Address),
    units_in_sale = n(),
    average_price = first(adjusted_price),
    .groups = "drop"
  ) |>
  filter(units_in_sale > 1) |>
  arrange(desc(`units_in_sale`))
head(multi_sales_collapsed, 15)
```
The before and after histograms of the Sale Price also demonstrate the effect of the data cleaning. The second histogram avoids the outlier issues that the original dataset has, namely the huge amount of properties with a sale price of 0 that throws off the scale of the chart, and the two large package sales of $14.2m and $20.5m that has a similar effect. The largest package sales can be seen in the dataframe above. Note that the cleaned histogram sees large spikes in count at $109k and $98k due to the two largest package sales that include 188 and 144 units in sale respectively. 

```{r}
ggplot(data = ames, aes(x = `TotalLivingArea (sf)`, y = `Sale Price`)) +
  geom_point() + 
  geom_smooth(method = "lm")

ggplot(data = ames_clean, aes(x = `TotalLivingArea (sf)`, y = `adjusted_price`)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

The end result is that we can observe a positive correlation between TotalLivingArea and adjusted_price in the scatterplot with cleaned data, as expected. As a property gets bigger, it's price should increase as it's more desirable. The scatterplots using both the original dataset and the cleaned dataset are provided as reference.



Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.
