---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Step 1 Result:
```{R}
library(classdata)
library(tidyverse)
ames = ames
View(ames)
```

```{r}
head(ames)
```

```{r}
names(ames)
```


```{r}
glimpse(ames)
```


```{r}
str(ames)
```

As a team, we inspected the ames dataset using glimpse(), names(), and str(). We found that the dataset contains 16 variables: Parcel ID, Address, Style, Occupancy, Sale Date, Sale Price, Multi Sale, YearBuilt, Acres, TotalLivingArea (sf), Bedrooms, FinishedBsmtArea (sf), LotArea(sf), AC, FirePlace, and Neighborhood. These variables include character variables such as identifiers and labels, factor variables such as housing style and neighborhood, one date variable (Sale Date), and several numeric variables such as Sale Price, YearBuilt, lot size, and living area. Based on the variable names and sample values, the dataset records property sale information, structural features, lot size, and location details for residential properties in Ames. We would expect identifiers and addresses to mostly appear as text, categorical variables to come from a limited set of repeated categories, sale dates to be valid dates, and numeric variables such as sale price, acreage, living area, and lot area to be nonnegative values. We also noticed that some variables contain missing values (NA), which may need to be considered during later analysis.





Step 2 Result:

```{r}
ames$`Sale Price`
```
Sale Price is the main variable in the dataset. It's the main variable we want to predict with given unseen data. Most other variables in the dataset will be directly correlated with the sale price of the property. 








Step 3 Result:











Step 4 Result:
Nicolas Result:







Nicole Result:






Scott Result:







Shashank Result:






Pablo Result:

```{r}

```





Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.
