---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Step 1 Result:
```{R}
library(classdata)
library(tidyverse)
ames = ames
View(ames)
```

```{r}
head(ames)
```

```{r}
names(ames)
```


```{r}
glimpse(ames)
```


```{r}
str(ames)
```







Step 2 Result:










Step 3 Result:
```{R}
price <-ames$`Sale Price`
range(ames[["Sale Price"]], na.rm = FALSE)

library(ggplot2)
ggplot(ames, aes(x = price)) +
  geom_histogram(binwidth = 1000000, color = "blue", fill = "white") +
  labs(title = "Sales Price",
       x = "Price ($)",
       y = "Frequency")


Q1 <- quantile(ames$`Sale Price`, 0.25, na.rm = TRUE)
Q3 <- quantile(ames$`Sale Price`, 0.75, na.rm = TRUE)
IQR_value <- IQR(ames$`Sale Price`, na.rm = TRUE)

lower_bound <- Q1 - 1.5 * IQR_value
upper_bound <- Q3 + 1.5 * IQR_value

ames_no_outliers <- ames[
  ames$`Sale Price` >= lower_bound &
  ames$`Sale Price` <= upper_bound,
]

ggplot(ames_no_outliers, aes(x = `Sale Price`)) +
  geom_histogram(binwidth = 100000, color = "blue", fill = "white") +
  labs(title = "Sales Price",
       x = "Price ($)",
       y = "Frequency")
```










Step 4 Result:
Nicolas Result: acres







Nicole Result:






Scott Result:







Shashank Result:






Pablo Result:







Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.
